### 2014-03-30 trunk/paper_enahpc14/paper.tex
updates introduction and bibliography

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% This is a general template file for the LaTeX package SVJour3
% for Springer journals.          Springer Heidelberg 2010/09/16
%
% Copy it to a new file with a new name and use it as the basis
% for your article. Delete % signs as needed.
%
% This template includes a few options for different layouts and
% content for various journals. Please consult a previous issue of
% your journal as needed.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
\RequirePackage{fix-cm}
%
\documentclass[twocolumn]{svjour3}           % twocolumn
%
\smartqed % flush right qed marks, e.g. at end of proof
%
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage[round, sort&compress, numbers]{natbib}
%
\DeclareRobustCommand\IPCClongname{}
%
% please place your own definitions here and don't use \def but
% \newcommand{}{}
%
% Insert the name of "your journal" with
% \journalname{myjournal}
%
\begin{document}

\title{Evaluating the performance  and energy efficiency of COSMO-ART,
  a fully online coupled model  system composed of a numerical weather
  forecast model  and a  chemical transport model.   \thanks{Grants or
    other notes  about the  article that should  go on the  front page
    should be placed here.}}
%\subtitle{Write here subtitle}

%\titlerunning{Short form of title} % if too long for running head

\author{J.Charles \and W.Sawyer \and M.Dolz \and C.Malossi}

%\authorrunning{Short form of author list} % if too long for running
%head

\institute{Dr. Joseph Charles \at Swiss National Supercomputing Centre
  (CSCS)\\ CH-6900  Lugano, Switzerland  \\ Tel: +41  (0) 91  610 8216
  \\ \email{joseph.charles@cscs.ch} \and Dr.  William Sawyer \at Swiss
  National Supercomputing Centre  (CSCS)\\ CH-6900 Lugano, Switzerland
  \\ Tel:  +41 (0) 91 610 8229  \\ \email{william.sawyer@cscs.ch} \and
  Dr.  Manuel Dolz \at University of Hamburg \\ 20148 Hamburg, Germany
  \\         Tel:         +49         (0)        40         460094-404
  \\ \email{manuel.dolz@informatik.uni-hamburg.de}  \and Dr. Cristiano
  Malossi  \at  IBM  Zurich  Research Laboratory  \\  CH-8000  Zurich,
  Switzerland     \\     Tel:      +41     (0)     44     724     8616
  \\ \email{ACM@zurich.ibm.com}}

\date{Received: date / Accepted: date} % The correct dates will be
                                       % entered by the editor

\maketitle

\begin{abstract}
  In this paper we present  COSMO-ART, an extension of the operational
  weather  forecast  model  of   the  German  Weather  Service  (DWD),
  developed for  the evaluation of the interactions  of reactive gases
  and aerosol particles  with the state of atmosphere  at the regional
  scale. It  includes secondary aerosols,  directly emitted components
  like  soot,  mineral  dust,  sea  salt and  biological  material  as
  pollen. Processes such  as emissions, coagulation, condensation, dry
  deposition,  wet removal,  and sedimentation  of aerosols  are taken
  into account.   The overall performance  of this application  on HPC
  systems is  analysed by a  profiling and tracing study  to determine
  hotspots  and  identify   critical  paths.   Moreover,  we  describe
  measurement devices and energy-aware techniques employed to evaluate
  the  energy  footprint of  the  considered  application  and to  get
  detailed  insights about  power bottlenecks.   Our motivation  is to
  improve  corresponding  code sections  to  sustain high  performance
  while  minimizing energy-to-solution.   This preliminary  study sets
  the basis of broader  considerations to tackle challenges related to
  energy efficient high performance  computing in the framework of the
  Exa2Green project (\url{http://exa2green.eu/}).

\keywords{High performance computing  \and Energy-aware computing \and
  Green computing  \and Numerical weather  prediction \and Atmospheric
  chemistry  \and  Aerosols  modelling  \and  Profiling  methods  \and
  Benchmark analysis \and COSMO-ART coupled model}
% \PACS{PACS code1 \and PACS code2 \and more}
% \subclass{MSC code1 \and MSC code2 \and more}
\end{abstract}

\section{Introduction}
\label{intro}
\input{Introduction.tex}

\section{COSMO coupled with ART extension}
\label{sec:1}
\subsection{Model description and set-up}
\label{subsec:1}
\input{Sec1:COSMO-ART.tex}

\section{Hardware description}
\label{sec:2}

\section{Power measurement systems}
\label{sec:3}

\section{Benchmark results}
\label{sec:4}

%\paragraph{Paragraph headings} Use paragraph headings as needed.

%\begin{figure*}
%  \includegraphics[width=0.4\textwidth]{Figs/earth.eps}
%  \caption{Please write your figure caption here}
%  \label{fig:1}
%\end{figure*}

%\begin{table}
%  \caption{Please write your table caption here}
%  \label{tab:1}
%  \begin{tabular}{lll}
%    \hline\noalign{\smallskip}
%    first & second & third  \\
%    \noalign{\smallskip}\hline\noalign{\smallskip}
%    number & number & number \\
%    number & number & number \\
%    \noalign{\smallskip}\hline
%  \end{tabular}
%\end{table}

\begin{acknowledgements}
This  research  was  supported   by  the  Exa2Green  research  project
co-funded  under the  EU  7th Framework  Program  Future and  Emerging
Technologies (FET) Proactive Initiative: Minimising Energy Consumption
of Computing to  the Limit (MINECC).  The authors  would like to thank
the  High  Performance  and  High  Productivity  Computing  Initiative
(\url{www.hp2c.ch}).)
\end{acknowledgements}

\DeclareRobustCommand\IPCClongname{ - Intergovernmental Panel on Climate Change}

\bibliographystyle{plainnat}
\bibliography{\jobname}

\end{document}