### 2014-04-01 trunk/paper_enahpc14/Sec2:Power-measurement.tex
adds missing files from previous commit

\subsection{Power-Performance measurement framework}
\label{subsec:3.1}
To assess the  performance and the energy efficiency  of COSMO-ART, we
employ   a    version   of   the    integrated   framework   presented
in~\cite{energy13}  that  works in  combination  with VampirTrace  and
Vampir,   which  are   profiling/tracing   and  visualization   tools,
respectively.
%The left part of the \vref{fig:Lustre} offers a graphical representation of
%the Lustre architecture; the right depicts the tracing and profiling framework.
To  use our  approach,  COSMO-ART is  compiled  using the  VampirTrace
compiler wrappers, which automatically  instrument the Fortran code of
the  model. Next,  COSMO-ART is  run  on the  nodes, thus  dissipating
certain  amount of  power.  The  server nodes  are connected  to power
measurement  devices that  account for  the  dissipated power/consumed
energy and  send the power data  to the tracing  server.  The attached
VampirTrace \pmlib plugin employs the client API that sends start/stop
primitives in order to gather captured data by the wattmeters onto the
tracing server,  where an  instance of the  \pmlib server  is running.
Once COSMO-ART run is finished, the VampirTrace \pmlib plugin receives
the  power   data  from  the  tracing   server.   The  instrumentation
post-process  generates the  performance  trace files  and the  \pmlib
plugin inserts the power data into them.

In  addition  to the  power  measurements,  we  also account  for  the
resource utilization values  of the nodes: CPU load,  memory usage and
storage device utilization. % and network utilization.

We  run special  \pmlib  server  instances on  the  server nodes  that
retrieve these  values from the \texttt{proc}  file system (leveraging
the  \texttt{psutil} Python library).   Thus, \pmlib  plugin instances
running  with the  instrumented  application connect  with the  \pmlib
servers.   Finally,   using  the   Vampir   visualization  tool,   the
power-performance traces  can be easily  analyzed through a  series of
plots and statistics.