### 2014-04-12 paper_enahpc14/Introduction.tex
updates Introduction; refreshes bibliography

While anthropogenic greenhouse gas emissions are driving unprecedented
major  climate  changes  since   the  mid  20th  Century,  one  factor
overwhelmingly  affects  the  uncertainty  in  defining  human-induced
radiative  forcing:  the  effects   of  aerosols  in  the  atmosphere.
Although they  are not considered  a heat-trapping greenhouse  gas and
have  shorter  atmospheric lifetimes,  they  significantly modify  the
global   radiation   budget   \citep{IPCC-2013}.   Enhanced   aerosols
concentrations impact the climate  system by scattering and absorption
of   solar   radiation,    thereby   exerting   a   direct   radiative
forcing  \citep{Liao-2005,  Bangert-2012,  Lundgren-2013}, or  by  the
modification of  cloud properties, cloud fraction  and surface albedo,
causing  a negative  indirect  radiative forcing  \citep{Haywood-2000,
Lohmann-2005,     VandenHeever-2011,     Rosenfeld-2013}.      Despite
considerable  progress in  global aerosol  modelling \citep{Mann-2013}
and    measurement-based    assessments   \citep{Myhre-2009},    large
uncertainties    remain    in    current    estimates    of    aerosol
forcing   \citep{Myhre-2013,    IPCC-2013,   Lee-2013,   Randles-2013,
Rosenfeld-2013, Sherwood-2013, Stier-2013}.

Hence to  improve our understanding of  aerosol-cloud interactions and
reduce  uncertainties  in  aerosol  radiative forcing  estimates,  the
research  community  is making  a  concerted  international effort  to
represent  the  underlying physical,  chemical  and aerosol  dynamical
processes through numerical chemical  transport models (CTMs), such as
ART (Aerosols  and Reactive Trace  gases), developed at  the Karlsruhe
Institute   of  Technology   (KIT)   \citep{Vogel-2009,  Bangert-2011,
Knote-2013}. This  regional scale model system is  online coupled with
the  operational weather  forecast  model COSMO  \citep{Baldauf-2011},
jointly developed by a consortium of European weather centers and used
in the climate version (COSMO-CLM)  by a wide research community.

The extended  COSMO-ART model  is computationally much  more demanding
than the  COSMO core  since a large  number of additional  tracers and
processes  have  to  be  considered.   Thus this  model  is  currently
severely-limited in  terms of applicability and expensive  in terms of
energy consumption.   Although COSMO has recently been  ported to GPUs
within  the framework of  the High  Performance and  High Productivity
Computing  (HP2C)  Initiative to  optimize  it  for computational  and
energy efficiency (\url{http://www.hp2c.ch/}), significant investments
in  ART  are still  required  to  take it  to  a  similar level.   The
efficiency of COSMO-ART is being addressed in the EU Exa2Green project
(\url{http://exa2green.eu/}). The  ultimate goal of the  project is to
deliver a  prototype code, which  provides an energy efficiency  of at
least five times of the  baseline value.  Such an implementation would
allow  the  community  to  investigate critical  questions  at  higher
resolution  and   over  longer  periods,   at  reduced  cost   to  the
environment.

Some  related work  will  be discussed  in  Sec.~\ref{sec:1}. Then  in
Sec.~\ref{sec:2}, we briefly  introduce the regional atmospheric model
COSMO-ART \citep{Vogel-2009}  and specify its  technical setup related
to  the  investigated   performance  and  energy  evaluation  methods.
Sec.~\ref{sec:3} describes different power measurement equipments used
for  conducting  this  benchmarking  study and  presents  energy-aware
techniques  employed   to  determine  the  energy   footprint  of  the
considered  application.  Sec.~\ref{sec:4}  gathers  benchmark results
from  the Exa2Green  project and  highlights areas  where improvements
will be  necessary for the subsequent  baseline refactoring.  Finally,
we  conclude with  practical guidelines  to identifying  energy saving
opportunities  and give  directions  for further  developments of  the
modeling system in Sec.~\ref{concl}.