### 2014-04-15 paper_enahpc14/Sec2:COSMO-ART.tex
updates Section 2: "The COSMO-ART model system"

\subsection{Model Description}
\label{subsec:1.1}
The  model  system COSMO-ART  described  in  \citep{Vogel-2009}, is  a
regional  to  continental scale  model  coupled  online  to the  COSMO
numerical weather  prediction and climate  model \citep{Baldauf-2011}.
It  incorporates  sophisticated  modules  for  gaseous  chemistry  and
aerosol dynamics  and allows the online calculation  of reactive trace
substances and their interaction  with the atmosphere.  Detailed model
description   can  be   found   in  \citep{Bangert-2012,   Knote-2011,
  Knote-2013}.

\subsection{Model Setup}
\label{subsec:1.2}
Establishing effective energy performance benchmarking of a code under
intense development such as COSMO-ART is a challenging task because of
the  absolute necessity that  results must  be reproducible  within an
expected variance for the duration of the Exa2Green project. To define
a baseline,  it was necessary  to find a run-configuration  capable of
being  recreated in  lal subsequent  versions.  Here  we give  a brief
overview of the model setup  for our performance and energy efficiency
evaluation.

Three-dimensional simulations are performed over large parts of Europe
and the Mediterranean Sea for April 13th 2010 in order to benefit from
around  half  day  of  sun  exposure and  therefore  ensure  a  proper
activation of the chemistry cycle.  The calculation domain corresponds
to the CORDEX-EU-44 domain and is covered by a grid of $222\times 216$
points with  a horizontal resolution of $0.22\,^{\circ}$,  i.e.  50 km
in  both directions and  40 vertical  layers.  These  24-hour forecast
simulations  are not  preceded  by  a spin-up  phase  to build-up  the
simulated gaseous and  aerosol concentrations.  This condition entails
that at the  model initialization there is a  period of adjustment for
the model  to reach its equilibrium  state and minimize  the effect of
initial conditions for gases and aerosols on the model predictions.

The meteorological initial and  bounday conditions are obtained by the
the ECMWF  global spectral model IFS  with an update  frequency of 3h.
Boundary data  for gas-phase species are taken  from IFS-MOZART output
at 6h temporal resolution. The  model setup incorporates 34 2-d and 45
3-d  fields  to be  written  out  every hour  and  is  devoid of  data
assimilation methods.

The  considered  COSMO-ART model  system  is  configured  with a  semi
Lagrangian horizontal advection scheme with tricubic interpolation and
selective filling  diffusion option in combination  with the dynamical
core  using Runge-Kutta  time  stepping \citep{COSMO-PartI-2011}.   It
also  makes use  of  the  Kinetic PreProcessor  solver  (KPP) for  the
resolution  of atmospheric  chemistry ordinary  differential equations
\citep{Damian-2002}.   Concerning the modelling  of wet  deposition in
aerosols, the  baseline has only indirect cloud  feedbacks but doesn't
include  in-cloud  scavenging  (rainout)  and  below-cloud  scavenging
(washout)  yet.   Amongst  physical  parameterisations,  precipitation
formation is performed by a two-moment cloud microphysics.