### 2014-04-16 paper_enahpc14/Sec3:Measurement-framework.tex
updates "Power-performance measurement framework" & "Evaluation" sections; update bibliography

In this section,  we present two different frameworks  deployed on HPC
systems  to  measure the  power  consumption  and  performance of  the
baseline execution.

\subsection{Framework description}
\label{subsec:3.1}

\subsubsection{E3METER metering products}
Supercomputer  clusters considered  for our  experiments at  the Swiss
National Supercomputing Center of  ETH Zurich (CSCS) are equipped with
E3METER Intelligent  Power Strips (IPS)  and Monitors (IPM)  which are
high   quality   electricity  meters   released   by  Riedo   Networks
(\url{http://riedonetworks.com/}),  that  enable  to monitor  and  log
power  consumption of  the  IT infrastructure  as  well as  constantly
analyze  line  voltage,  current,  power-factor,  frequency  with  1\%
accuracy.   Using reliable  narrowband  powerline communication  (PLC)
technology,  all metering  and power  quality data  from each  IPS are
centrally collected by the E3METER Data Concentrator, via the existing
power cables thus  avoiding the need for extra  cabling.  This data is
made  available  via SNMP,  HTTP,  TELNET  through  the built-in  Fast
Ethernet  port.   Time  synchronization  is guaranteed  by  using  NTP
servers.   Measured data  is accessed  through the  open  source Cacti
software including  the E3METER  Cacti Plugin to  scan the  entire PLC
network and monitor  in real-time the power usage  of individual rack,
recorded in 5 minutes interval periods.

\subsubsection{University of Hamburg}
To assess the  performance and the energy efficiency  of COSMO-ART, we
employ   a    version   of   the    integrated   framework   presented
in~\cite{energy13}  that  works in  combination  with Extrae  and
Paraver,   which  are   profiling/tracing   and  visualization   tools,
respectively.
%The left part of the \vref{fig:Lustre} offers a graphical representation of
%the Lustre architecture; the right depicts the tracing and profiling framework.
To  use our  approach,  COSMO-ART is  compiled  using the  Extrae
compiler wrappers, which automatically  instrument the Fortran code of
the  model. Next,  COSMO-ART is  run  on the  nodes, thus  dissipating
certain  amount of  power.  These  nodes are connected  to power 
measurement  devices that  account for  the dissipated power/consumed 
energy and  send the power data  to the tracing  server. The client, 
meanwhile, sends start/stop primitives in order to gather captured data
by the wattmeters onto the tracing server, where an  instance of the 
\pmlib server  is running.

Once COSMO-ART run is finished, a file containing the power profile is 
created using the  power   data  received from  the  tracing   server and
the  instrumentation post-processing generates  the performance trace files.
All these files are combined then in Paraver which allows us to visualize 
the performance trace and the power profile of COSMO-ART all together.

%Once COSMO-ART run is finished, the VampirTrace \pmlib plugin receives
%the  power   data  from  the  tracing   server.   The  instrumentation
%post-processing generates  the performance trace files  and the \pmlib
%plugin inserts the power data into them.

%In  addition  to the  power  measurements,  we  also account  for  the
%resource utilization values  of the nodes: CPU load,  memory usage and
%storage device utilization. % and network utilization.

%We  run special  \pmlib  server  instances on  the  server nodes  that
%retrieve these  values from the \texttt{proc}  file system (leveraging
%the  \texttt{psutil} Python library).   Thus, \pmlib  plugin instances
%running  with the  instrumented  application connect  with the  \pmlib
%servers.   Finally,   using  the   Vampir   visualization  tool,   the
%power-performance traces  can be easily  analyzed through a  series of
%plots and statistics.