### 2014-04-19 trunk/paper_enahpc14/Sec5:Supercomputers.tex

In this section we provide a brief description of the HPC supercomputers
used to obtain the COSMO-ART baseline. For our power measurement we have chosen state-of-the-art Intel and IBM platforms;
both architectures could be recreated or found in an identical configuration
beyond the lifetime of the project.  Given that the baseline benchmark
can be reproduced  within an expected variance, and  that the baseline
run configuration  can be used in  all future versions of  the code, a
fair comparison  will be made  between the baseline and  the future milestone
versions   of    COSMO-ART. In addition, a   complementary energy-to-solution  benchmarking  comparison  is  carried out  on  the
``Pilatus'' cluster based  on Intel's previous-generation Sandy Bridge
processors, conventional  in HPC  systems and known  to be  more power
consuming.   Finally a  full tracing  experiment is  conducted  on the
``Tintorrum''  cluster using  an integrated  framework, to  capture an
overall  power profile  at a  much finer  resolution and  to  get more
insights  into  the  MPI  blocking  and polling  influences  on  power
savings.

% In  this section,  we  describe a  performance  and energy  efficiency
% evaluation  of  different  achitectures  when  running  the  COSMO-ART
% baseline.  We  start by  specifying our measurement  methodology along
% with the metrics used to  analyse the results on all platforms.  Then,
% we detail the environment  setup gathering the considered HPC systems,
% the  software  environment and  the  run  configuration.  Finally,  we
% discuss benchmark  results and  power-performance traces of  the model
% system.

% \subsection{Environment setup}
% \label{subsec:4.2}
% While hardware  platforms mature  and are replaced,  we have  chosen a
% state-of-the-art  Intel's  third-generation   Core  (aka  Ivy  Bridge)
% processing platform (called ``Monch'')  for our power measurements, as
% it  is slated  to stay  in service  without hardware  upgrade  for the
% duration  of  the Exa2Green  project.   In  principle  at least,  this
% architecture could be recreated or found in an identical configuration
% beyond the lifetime of the project.  Given that the baseline benchmark
% can be reproduced  within an expected variance, and  that the baseline
% run configuration  can be used in  all future versions of  the code, a
% fair comparison  will be made  between the baseline and  the milestone
% versions   of    COSMO-ART.    In   this    study,   a   complementary
% energy-to-solution  benchmarking  comparison  is  carried out  on  the
% ``Pilatus'' cluster based  on Intel's previous-generation Sandy Bridge
% processors, conventional  in HPC  systems and known  to be  more power
% consuming.   Finally a  full tracing  experiment is  conducted  on the
% ``Tintorrum''  cluster using  an integrated  framework, to  capture an
% overall  power profile  at a  much finer  resolution and  to  get more
% insights  into  the  MPI  blocking  and polling  influences  on  power
% savings.

\subsection{Monch (CSCS - ETH Zurich)}
Monch, which is installed  at the Swiss National Supercomputing Center
(CSCS) of ETH Zurich, is  a 10 rack NEC-provided and dual-socket Intel
Ivy Bridge-EP based cluster, utilised from people that are part of the
Swiss    Platform   for    Advanced   Scientific    Computing   (PASC,
\url{http://www.pasc-ch.org/}). It is composed of 312 standard compute
nodes, 24 large-memory compute nodes and 24 huge-memory compute nodes.
Each standard  compute node comprises  two Intel Ivy  Bridge Efficient
Performance (EP)  E5-2660 v2 ten-core processors operating  at 2.2 GHz
base  clock speed,  themselves connected  by a  high  speed InfiniBand
network based on Mellanox SX6036  managed FDR switches, with a 56 Gb/s
speed.  Each core  has 32 KB instruction and 32 KB  data L1 caches and
256 KB  of L2 cache. All  the 10 core share  a 25 MB L3  cache and the
platform has  32GB of  DDR3 1600 MHz  RAM. For  our energy-to-solution
benchmark, a  full rack  of Monch constituted  of 52  standard compute
nodes (monchc[029-080]) was considered.

\subsection{Pilatus (CSCS - ETH Zurich)}
Pilatus,  which  is installed  at  the  Swiss National  Supercomputing
Center (CSCS) of  ETH Zurich, is a dual-socket  eight-core Intel Sandy
Bridge EP based cluster used as Piz Daint pre-post processing cluster.
It is composed of 42  compute nodes and has 2 high-speed interconnects
based on FDR: the first is dedicated to the MPI traffic and the second
to  the storage  high speed  traffic.  The  2 login  nodes and  the 42
computes nodes consists in 11 twin-pair Intel E5-Series DALCO r2264i4t
2U  scalable compute modules.   Each module  contains 4  compute nodes
based on two  Intel Xeon E5-2670 processors operating  at 2.6 GHz base
clock speed,  themselves connected by a high  speed InfiniBand network
based  on  Mellanox  SX6036  managed  FDR switches,  with  a  56  Gb/s
speed. Each  core has 32 KB instruction  and 32 KB data  L1 caches and
256 KB  of L2 cache.  All  the 8 core share  a 20 MB L3  cache and the
platform has  64GB of  DDR3 1600 MHz  RAM. For  our energy-to-solution
benchmark, a full  rack of Pilatus constituted of  42 standard compute
nodes (pilatus[03-44]) was considered.

\subsection{Tintorrum (UJI)}
Tintorrum, which  is installed at the  Jaume I University  (UJI), is a
heterogeneous  cluster   composed  of  28  compute   nodes.   For  our
experiments only a subset of  these nodes were considered. This set is
composed  of 16 nodes,  each of  which includes  two Intel  Xeon E5645
hexa-core  processors  at  2.40   GHz  connected  via  Infiniband  QDR
(Mellanox MTS3600 switch).  Each core  has 32 KB instruction and 32 KB
data L1 caches and 256 KB L2 cache. The 6 cores share a 12 MB L3 cache
and the platform has 288 GB of DDR3 1333 MHz.

\subsection{Blue Gene/Q (IBM Research -- Zurich)}
The IBM Blue~Gene/Q~(BG/Q) supercomputer~\cite{IBMRedBookBGQHardware} is the
third-generation computer architecture in the Blue~Gene family. It is composed
by one or more compute racks, where each rack contains 1024 16-core compute cards,
with a peak performance of 209~TFLOPS per rack. Each compute card contains 16~IBM~Blue~Gene/Q
PowerPC~A2 core processors and 16 GB of memory. The chip has 16~cores (with
up to 4~threads per core) at~1.60~GHz, with 16~kB~L1 cache each. In addition,
it has 32~MB~L2 shared cache, and 2~controllers which provide access to 2~RAM banks
of 8~GB each, both at~1333~MHz. Moreover, each core has an L1~prefetcher that is shared by
the threads on the core. The network topology for Blue Gene/Q is a five-dimensional torus,
with direct links between the nearest neighbors in the $\pm$A, $\pm$B, $\pm$C, $\pm$D, and $\pm$E
directions and 40~GBps bandwidth.