### 2014-04-22 paper_enahpc14/Introduction.tex
commits Will's slight corrections

While anthropogenic greenhouse gas emissions are driving unprecedented
major  climate  changes  since   the  mid  20th  Century,  one  factor
overwhelmingly  affects  the  uncertainty  in  defining  human-induced
radiative  forcing:  the  effects   of  aerosols  in  the  atmosphere.
Although they  are not considered  a heat-trapping greenhouse  gas and
have  shorter  atmospheric lifetimes,  they  significantly modify  the
global   radiation   budget   \citep{IPCC-2013}.   Enhanced   aerosols
concentrations impact the climate  system by scattering and absorption
of   solar   radiation,    thereby   exerting   a   direct   radiative
forcing  \citep{Liao-2005,  Bangert-2012,  Lundgren-2013}; or  by  the
modification of  cloud properties, cloud fraction  and surface albedo,
causing  a negative  indirect  radiative forcing  \citep{Haywood-2000,
Lohmann-2005,     VandenHeever-2011,     Rosenfeld-2013}.      Despite
considerable  progress in  global aerosol  modelling \citep{Mann-2013}
and    measurement-based    assessments   \citep{Myhre-2009},    large
uncertainties  remain  in   current  estimates  of  aerosol  radiative
forcing   \citep{Myhre-2013,    IPCC-2013,   Lee-2013,   Randles-2013,
Rosenfeld-2013, Sherwood-2013, Stier-2013}.

Hence to  improve our understanding of  aerosol-cloud interactions and
reduce  these  uncertainties,  the  research  community  is  making  a
concerted international  effort to represent  the underlying physical,
chemical  and aerosol dynamical  processes through  numerical chemical
transport  models (CTMs)  such  as ART  (Aerosols  and Reactive  Trace
gases),   developed   at  the   Karlsruhe   Institute  of   Technology
(KIT)  \citep{Vogel-2009,  Bangert-2011,  Knote-2013}.  This  regional
scale model  system is coupled  with the operational  weather forecast
model COSMO \citep{Baldauf-2011}, jointly developed by a consortium of
European weather  centers and used in the  climate version (COSMO-CLM)
by a wide research community.

The extended  COSMO-ART model  is computationally much  more demanding
than COSMO  since a large  number of additional tracers  and processes
have to  be considered.  Thus the model  is currently severely-limited
in  terms   of  applicability  and   expensive  in  terms   of  energy
consumption.    Although   COSMO   has   recently   been   ported   to
GPUs  \citep{Gysi-2014, Lapillonne-2014} within  the framework  of the
High Performance and High  Productivity Computing (HP2C) Initiative to
optimise    it    for     computational    and    energy    efficiency
(\url{http://www.hp2c.ch/}), significant investments  in ART are still
required to take  it to a similar level.   The efficiency of COSMO-ART
is     being    addressed    in     the    EU     Exa2Green    project
(\url{http://exa2green.eu/}). The  ultimate goal of the  project is to
deliver a  prototype code, which  provides an energy efficiency  of at
least five times of the  baseline value.  Such an implementation would
allow  the  community  to  investigate critical  questions  at  higher
resolution  and   over  longer  periods,   at  reduced  cost   to  the
environment.

This  work is organized  as follows:  in Sec.~\ref{sec:1}  we describe
related  work,  then in  Sec.~\ref{sec:2},  we  briefly introduce  the
regional atmospheric  model COSMO-ART and specify  its technical setup
related to the investigated performance and energy evaluation methods.
Sec.~\ref{sec:5} describes the HPC  supercomputers used to compute the
baseline.  Sec.~\ref{sec:3}   describes  different  power  measurement
equipment  used for  conducting this  benchmarking study  and presents
energy-aware techniques employed to  determine the energy footprint of
the   considered  application.   Sec.~\ref{sec:4}   gathers  benchmark
results  from  the  Exa2Green   project  and  highlights  areas  where
improvements   will   be  necessary   for   the  subsequent   baseline
refactoring.   Finally,  we  conclude  with  practical  guidelines  to
identifying energy saving opportunities  and give an outlook to future
research in Sec.~\ref{concl}.