### 2014-04-23 paper_enahpc14/Sec1:Related.tex
adds minor changes; updates paper.bib; includes references for related work

\begin{itemize}
\item Application centric energy-efficiency study of distributed
  multi-core and hybrid {CPU-GPU} systems \citep{Cumming-2014}
\item Stella: A domain-specific language and tool for structured grid
  methods \citep{Gysi-2014}
\item Using compiler directives to port large scientific applications
  to GPUs: An example from atmospherci science \citep{Lapillonne-2014}
\item An analysis of energy-optimized lattice-Boltzmann CFD
  simulations from the chip to the highly parallel level
  \citep{Wittmann-2013}
\item Energy efficiency vs. performance of the numerical solution of
  PDEs: An application study on a low-power ARM-based cluster
  \citep{Goddeke-2013}
\item Analysis of Computing and Energy Performance of Multicore, NUMA,
  and Manycore Platforms for an Irregular Application
  \citep{Castro-2013}
\item Exploring performance and power properties of modern multi-core
  chips via simple machine models \citep{Hager-2013}
\end{itemize}