### 2014-04-25 paper_enahpc14/Conclusion.tex
adds conclusion; modifies table

We  have   presented  a  methodology  for   comparing  performance  of
COSMO-ART,  a  regional  weather  forecast  model  augmented  for  the
interactions of  reactive gases and aerosol  particles.  The resulting
benchmarks illustrate  that the  best time-to-solution does  not imply
the best energy-to-solution: an Intel Sandybridge (2.6 GHz) system has
lower  time-to-solution but  higher energy-to-solution  than  an Intel
Ivybridge  (2.2  GHz) system,  although  the  two  metrics are  indeed
strongly  correlated.   A   tool  utilising  Paraver/Extrae  with  new
extensions  for power  consumption  was successfully  applied to  this
non-trivial application.  The  resulting profiles indicate that simple
changes, such as making use of an MPI version based on blocking rather
than polling, can reduce power consumption significantly.

This  reproducible  benchmark provides  a  baseline  for ongoing  work
package within  the EU-funded Exa2Green to  minimise power consumption
of COSMO-ART. Profiling  has given us insight into  the most expensive
code  components,  which are  now  being  altered  to utilise  revised
algorithms.  The  results of these  optimisations will be  reported in
future publications.