### 2014-05-05 trunk/paper_enahpc14/Sec3:Environment-setup.tex
cleans up latest version; improves format

\subsection{Hardware platforms}
\label{subsec:3.1}

Benchmarks presented  in this work  were performed on  three different
platforms  and can  be reproduced  within an  expected  variance, thus
allowing a  fair comparison between the baseline  and future milestone
versions of \cosmoart.

\begin{itemize}
\item \monch is  a 10-rack NEC-provided clus\-ter at  CSCS composed of
  312 compute nodes. 
  % and  utilized  by  members  of  the  Swiss  Platform  for  Advanced
  % Scientific Computing (PASC) project~\cite{PASC}.
  Each node is  comprised of two Intel Ivy  Bridge EP E5-2660 ten-core
  processors  operating  at 2.2\,GHz,  equipped  with  32\,GB of  DDR3
  1600\,MHz RAM  and connected via InfiniBand Mellanox  SX6036 and FDR
  switches. For our  experiments, a full rack constituted  of 52 nodes
  was used. 
  % \monch is  slated to stay  in service without hardware  upgrade for
  % the  duration  of  the  Exa2Green project,  allowing  an  identical
  % configuration  to  be  recreated  for  future  assessments  of  the
  % baseline.

\item \pilat is  a dual-socket eight-core Intel Sandy  Bridge EP based
  cluster at CSCS composed of 42 compute nodes.  
  % and used as Piz Daint  pre-post processing cluster The 2 login and
  % 42  compute  nodes  are   consolidated  into  11  twin-pair  Intel
  % E5-Series  DALCO  r2264i4t   2U  scalable  compute  modules,  each
  % containing 4 compute nodes
  Each node  is comprises two Intel Xeon  E5-2670 processors operating
  at 2.6\,GHz equipped with 64\,GB of DDR3 1600\,MHz RAM and connected
  via  InfiniBand   Mellanox  SX6036   and  FDR  switches.    For  our
  experiments, a full  rack constituted of 42 compute  nodes was used.
  \pilat   is   based   on   Intel's   second-generation   processors,
  conventional in  HPC systems  and known to  be more  power consuming
  than its Ivy Bridge successor.

\item \tinto is  a heterogeneous cluster composed of  28 compute nodes
  at UJI.  For our experiments only  a subset of  16 homogeneous nodes
  was  considered. Each  node  is  composed of  two  Intel Xeon  E5645
  hexa-core processors  running at  2.4\,GHz, equipped with  24\,GB of
  DDR3  1333\,MHz and  connected via  InfiniBand QDR  with  a Mellanox
  MTS\-3600 switch.
\end{itemize}

\subsection{Power measurement framework}
\label{subsec:3.3}

We present here  the two different frameworks deployed  on the testbed
HPC  systems  to measure  performance,  power  dissipation and  energy
consumption of the baseline execution of \cosmoart.

\subsubsection{E3METER metering framework}

Supercomputer clusters  considered for our experiments at  CSCS of ETH
Zurich are  equipped with E3METER  Intelligent Power Strips  (IPS) and
Monitors (IPM)  which are high quality electricity  meters released by
Riedo Networks,  that enable to  monitor and log power  dissipation of
the  IT infrastructure  as well  as constantly  analyze  line voltage,
current, power-factor, frequency  with 1\,\% accuracy.  Using reliable
narrowband power-line communication (PLC) technology, all metering and
power  quality data  from  each  IPS are  centrally  collected by  the
E3METER Data Concentrator, via the existing power cables thus avoiding
the need  for extra  cabling.  This data  is made available  via SNMP,
HTTP  and  TELNET  through  the  built-in Fast  Ethernet  port.   Time
synchronization is guaranteed by  using NTP servers.  Measured data is
accessed through the open  source Cacti software including the E3METER
Cacti Plugin to  scan the entire PLC network  and monitor in real-time
the power  usage of  individual rack, recorded  in 5  minutes interval
periods.

\subsubsection{Power-performance tracing framework}

For  experiments  at UJI  on  \tinto, we  employed  a  version of  the
integrated  framework presented  in~\cite{energy13}  that employs  the
\pmlib  library in  combination  with Extrae  and  Paraver, which  are
profiling/tracing and visualization tools, respectively.

To employ our  framework, \cosmoart is first compiled  and linked with
the  Extrae instrumentation  library, which  automatically instruments
the Fortran~90  code of the model,  e.g., to record  MPI calls.  Next,
\cosmoart  is run  on the  nodes, thus  dissipating certain  amount of
power. The  nodes are connected  to APC 8653 Power  Distribution Units
(PDUs) that account  for the dissipated power with  a sampling rate of
1\,Hz. The power data captured by  these PDUs is collected by a \pmlib
tracing server running  on a separate machine. The  client, running on
the target nodes, sends start/stop  primitives in order to collect the
power  data during  the \cosmoart  run. Once  the run  is  finished, a
power-performance  trace file  containing the  power data  is received
from  the  tracing  server  and the  instrumentation  post-process  is
generated.  A  merged trace  file  is  finally  loaded in  Paraver  to
visualize  and correlate  the  performance trace  and  with the  power
profile  generated by \cosmoart.  This procedure  allows us  to easily
detect power bottlenecks, basically constituted by inefficient waiting
methods.

\subsection{Software environment}
\label{subsec:3.2}

The \cosmoart  baseline is a pure MPI-based  Fortran~90 code currently
running on distributed multi-core systems only.  The software stack on
the platforms  was controlled using the modules  framework which gives
an  easy and  flexible  mechanism to  access  to all  of the  provided
compilers, tools  and applications.  For our  initial benchmarking, we
opted for the GNU compiler GCC 4.8.1 on \monch and GCC 4.8.2 on \pilat
and  \tinto)  using  the -O3  compiler  flag  in  favor of  the  Intel
compilers 14.0.1, which  delivered inferior performance.  In addition,
we  installed the following  implementations of  MPI: MVAPICH2  1.9 on
\monch and \pilat, and OpenMPI 1.6.5  on \tinto. On the other hand, we
used HDF5  1.8.12 and NetCDF  (4.3.1) libraries for the  management of
extremely large and complex data collections.  All computes nodes have
an     operating    system     based     on    GNU/Linux     featuring
``2.6.32-358.11.1.el6'' kernel  in \monch, ``3.0.101-0.15-default'' in
\pilat and ``2.6.18-238.19.1.el5'' in \tinto.

A snapshot of the code, which includes, at least conceptually, all the
information needed  to reproduce the  energy-to-solution benchmarks of
\cosmoart, was produced  and run on 1040 cores  using 20~MPI tasks per
node on  \monch, on 1344 cores  using 16~MPI tasks per  node on \pilat
and  on  192  cores  using  12~MPI  tasks per  node  on  \tinto.   The
calculated region  was mapped to the participating  processors using a
2D-partitioning  strategy.  The  distribution  along the  $x$ and  $y$
coordinates  was   defined  accordingly  depending   on  the  selected
platform.  Besides,  as  this  version   does  not  make  use  of  the
traditional  GRIB  library, we  specified  $nprocio=0$  for GRIB  I/O.
Hyperthreading  was not  considered  in this  study  as it  previously
revealed that it always led to higher energy-to-solution.