### 2014-05-05 paper_enahpc14/Sec4:Experimental-results.tex
cleans up latest version; improves format

In this section, we present our measurement methodology along with the
metrics  considered  and  discuss about  the  performance-power-energy
trade-off of the model system on the different platforms stated.

\subsection{Measurement methodology}
\label{subsec:4.1}

We  approach  the  assessment  of  the energy  footprint  and  overall
performance    of    \cosmoart    with    two    important    metrics:
\textit{time-to-solution} (TTS) and \textit{energy-to-solution} (ETS).
TTS refers to  the total wall clock time  of the application execution
time. ETS  is the amount of  energy spent to  achieve results.  Energy
consumption is  assessed by sampling the power  during execution which
is then averaged and multiplied  by the TTS to determine ETS. Whenever
possible, multiple  production runs  were performed to  illustrate the
reproducibility  of   the  baseline,  and   quantify  the  significant
uncertainties in  the power measurement, as dictated  by the available
technology.

For  the   experiments  on  \tinto  platform,  we   also  analyze  the
contribution of the MPI library  to the energy consumption and how the
use  of  the  blocking  versus polling  message-passing  policies  can
potentially  render energy  savings. Specifically,  the  OpenMPI 1.6.5
library, features two operation modes, blocking and polling, which can
be selected before \cosmoart is launched. In the polling configuration
(the default mode) ``MPI''  continually polls the network interface to
check for  the completion  of an event  (e.g. send or  receive).  This
usually yields  low latency  but high CPU  utilization. Thus,  one can
expect that  this mode attains  the best performance, possibly  at the
cost a higher energy usage. In  the blocking mode, the CPU is yield to
other     processes/threads    if     there     are    no     incoming
messages~\cite{Castillo-2012}.

\subsection{Time-power-energy analysis}
\label{subsec:4.2}

Figures~\ref{fig:1} and \ref{fig:2}  account respectively for \monch's
Isola  E1 Rack 2  and \pilat'  Isola HD  total power  measurements for
1-day or 2-day  simulations. On the Intel Ivy  Bridge EP based cluster
(i.e.   \monch), the  1-day simulation  was issued  only twice  due to
usage restrictions. As  time resolution was set to  one update every 5
minutes  for  power  sampling,  the  averaged  power  dissipation  was
computed by  considering 6 values for  each single run.   On the Intel
Xeon E5 based cluster (i.e.   \pilat), the 1-day simulation was issued
four times and  a 2-day run only once.   Similarly, the averaged power
dissipation was computed by considering 4 values for each single 1-day
run  and  9 values  for  the  2-days  run. Corresponding  results  are
gathered in Table~\ref{tab:1}.

\begin{figure}[htbf]
  \centering
  \includegraphics[width=0.48\textwidth]{Figs/NRJ_benchmark_Monch.eps}
    \caption{Isola E1 Rack 2 total power of \monch.}
    \label{fig:1}
\end{figure}

\begin{figure}[htbf]
  \centering
  \includegraphics[width=0.48\textwidth]{Figs/NRJ_benchmark_Pilatus.eps}
    \caption{Isola HD total power dissipation of \pilat.}
    \label{fig:2}
\end{figure}

\begin{table}[htbf]
  \centering
    \caption{Averaged power dissipation (W) of the platforms.}
    \label{tab:1}
    \begin{tabular}{cccc}
      \hline\noalign{\smallskip}
      \textbf{\scriptsize{Xeon E5}} & \textbf{\scriptsize{Ivy Bridge}} & 
      \textbf{\scriptsize{Xeon  E5645}} & \textbf{\scriptsize{Xeon  E5645}} \\
      & & \textbf{\scriptsize{(Polling)}} & \textbf{\scriptsize{(Blocking)}} \\
      \noalign{\smallskip}\hline\noalign{\smallskip}
      18035.0 & 12622.5 & 3713.6 & 3651.8 \\ 
      \noalign{\smallskip}\hline
    \end{tabular}
  \end{table}

In  Figure~\ref{fig:3}, we  compare both  TTS (right  Y-axis)  and ETS
(left  Y-axis)  metrics  on  both  platforms.  As  expected,  Xeon  E5
outperforms Ivy  Bridge EP, being roughly 130\,\%  faster.  The reason
for that is two-fold: \emph{i}) it has higher clock frequency than Ivy
Bridge  (2.6\,GHz  against  2.2\,GHz),   and  \emph{ii})  it  aims  at
computing speed regardless to energy consumption.  In our experiments,
Ivy Bridge EP showed  the best energy-to-solution, reducing the energy
consumption of Xeon E5 by approximately 7\,\%.

Averaged TTS and ETS results  over 10 repetitions of 24h simulation on
\tinto are shown in  Figure~\ref{fig:4}. The corresponding results for
the  averaged total  power are  also shown  in  Table~\ref{tab:1}.  As
expected, the average total power dissipation by \cosmoart while using
the blocking MPI policy is about  2\,\% less than that consumed by the
polling  policy. This is  due to  the power-friendly  (blocking) waits
performed  by the  CPU cores  while using  the blocking  MPI  mode, in
contrast to  the polling  waits that take  place when the  polling MPI
mode is set. From the point of  view of the execution time, there is a
slight reduction  for the blocking MPI  mode. One can  expect that the
costs of the interruptions used in the blocking MPI policy can lead to
an  increase  of  the   execution  time,  however,  applications  with
extensive computing,  as e.g.  \cosmoart, may show  better performance
with the blocking  modes. This effect works in  favor with our results
since the energy consumed is also reduced (2\,\%) just by changing the
behavior of the OpenMPI library  to the blocking policy, i.e, with the
\texttt{--mca    mpi\_yield\_when\_idle}   parameter   set    in   the
\texttt{mpirun} command.

\begin{figure}[htbf]
  \centering
  \includegraphics[width=0.5\textwidth]{Figs/Time_E2S_COSMO-ART-0.eps}
  \caption{Time-to-solution and  energy-to-solution comparison between
    Xeon E5 and Ivy Bridge-EP architectures for a 24h simulation.}
  \label{fig:3}
\end{figure}

\begin{figure}[htbf]
  \centering
  \includegraphics[width=0.5\textwidth]{Figs/Time_E2S_COSMO-ART-1.eps}
  \vspace{-1cm}
  \caption{Mean time-to-solution and  energy-to-solution on \tinto for
    a 24h simulation using the  polling and blocking MPI policies. The
    95\,\%  confidence  intervals of  Student's  $t$ distribution  (10
    samples) clearly  indicates an improvement for  blocking policy in
    both TTS and ETS.}
  \label{fig:4}
\end{figure}

\begin{figure*}[ht]
  \centering
  \hspace{0.8cm} \scalebox{0.62}{\input{Figs/24hour.pstex_t}}
  \caption{24  hours  simulation power  trace  using  the polling  and
    blocking MPI policies on \tinto.}
  \label{fig:5}
\end{figure*}

\begin{figure*}[ht]
  \centering
  \scriptsize
  Simulation of 1 hour during midday (11h--12h) leveraging the polling MPI policy \\
  \scalebox{0.5}{\input{Figs/11_12_blq0_figure.pstex_t}} \\
  Simulation of 1 hour during midday (11h--12h) leveraging the blocking MPI policy \\
  \scalebox{0.5}{\input{Figs/11_12_blq1_figure.pstex_t}} \\
  1 hour simulation during midnight (23h--24h) leveraging the polling MPI policy \\
  \scalebox{0.5}{\input{Figs/23_24_blq0_figure.pstex_t}} \\
  1 hour simulation during midnight (23h--24h) leveraging the blocking MPI policy \\
  \scalebox{0.5}{\input{Figs/23_24_blq1_figure.pstex_t}} \\
  \begin{tabular}{rlp{-0.2cm}rlp{-0.2cm}rlp{-0.2cm}rlp{-0.2cm}rl}
& &  & &  & &  \\[-0.15cm]
Computation     & \multicolumn{1}{>{\columncolor[RGB]{  0,  , 255}}p{0.4cm}}{} & & 
Block. send     & \multicolumn{1}{>{\columncolor[RGB]{255,  0,174}}p{0.4cm}}{} & & 
Synchronization & \multicolumn{1}{>{\columncolor[RGB]{179,  0,  0}}p{0.4cm}}{} & & 
Wait/Wait all   & \multicolumn{1}{>{\columncolor[RGB]{235,255,  0}}p{0.4cm}}{} \\
& &  & &  & &  \\[-0.15cm]
Immediate recv. & \multicolumn{1}{>{\columncolor[RGB]{100,100,177}}p{0.4cm}}{} & &
Waiting a msg.  & \multicolumn{1}{>{\columncolor[RGB]{255,  0  ,0}}p{0.4cm}}{} & & 
Group comm.     & \multicolumn{1}{>{\columncolor[RGB]{255,144, 26}}p{0.4cm}}{} & &
Others          & \multicolumn{1}{>{\columncolor[RGB]{192,224,  0}}p{0.4cm}}{} \\
%I/O             & \multicolumn{1}{>{\columncolor[RGB]{172,174, 41}}p{0.4cm}}{} & &
%
%Idle            & \multicolumn{1}{>{\columncolor[RGB]{117,195,255}}p{0.4cm}}{} \\
\end{tabular}
  \caption{Power-performance   traces  during   midday   and  midnight
    leveraging the polling and blocking MPI policies on \tinto.}
  \label{fig:6}
\end{figure*}

\subsection{Power-performance profiling and tracing}
\label{subsec:4.3}

A full tracing  experiment is conducted on \tinto  in order to capture
an overall power profile at a much finer resolution using the attached
PDUs in each  node. First we run \cosmoart for  a 1-day simulation and
analyze the  power consumption.  Second we correlate  performance with
power  traces  of two  1-hour  time  frames  during the  day:  mid-day
(11h--12h) and midnight (23h--24h). All the experiments were performed
using 192 cores using 12 MPI processes per node on \tinto.

In order to obtain an initial overview of the power dissipation of the
system  model,  we  obtained  power  profiles  leveraging  our  \pmlib
power-measurement    framework   over    a    1-day   simulation    of
\cosmoart. These profiles, produced with both polling and blocking MPI
policies,  are depicted in  Figure~\ref{fig:5}. The  first observation
for the polling MPI policy is that the first and last hours of the day
dissipate slightly less power than  the mid-day hours. We relate these
variations  due   to  the  different   chemical  reactions,  aerosols,
radiations  and clouds  formation processes,  computed by  ART, taking
place during  daylight hours  but not during  night. In this  case the
peak power consumption is about  3821\,W. Looking at the power profile
using the blocking  MPI policy, we observe that  the 1-hour pattern is
repeated along  the 24-hours, favoring  the reduction of  power during
night  hours in  a  more noticeable  way  than that  with the  polling
policy.  The peak  power dissipation  during  day hours  of while  the
blocking  MPI policy  is leveraged  is  about 3768\,W,  which means  a
reduction of around  50\,W with respect to the  polling mode. Also, we
notice systematic  power drops for  each simulated hour.  We associate
them  to periods  in  which  the processes  were  blocked waiting  for
incoming messages  or synchronizations, thus  yielding the CPU  to low
usage. Our final observation is the reduction of the TTS by 2\,\% with
respect to the blocking MPI policy.

To  gain  more  insights   about  the  power-performance  behavior  of
\cosmoart,  we  obtained power-per\-for\-man\-ce  traces  to allow  us
correlate  the type  of  tasks  being executed  with  the total  power
dissipation  along  the   time  (see  Figure~\ref{fig:6}).   For  this
experiment we  limit the simulation  time to 1 hour  because \emph{i})
the  pattern of  the computation  performed by  \cosmoart  is repeated
hourly and, \emph{ii}) the weight  of a 1-day simulation trace can not
be easily handled using our  machines and memory available (each trace
file was about 45\,GB). To  face the problem we shrink the simulations
to 1 hour  and test a time  frame during midday (from 11h  to 12h) and
midnight  (from 23h  to  24h), in  which  \cosmoart perform  different
internal computations.  At  the same time we leverage  the two already
stated MPI policies: polling and blocking.

All  the  traces  in   Figure~\ref{fig:6}  are  divided  in  different
computation  phases. In first  phase all  the processes  perform group
communications      and       execute      MPI-receive      primitives
(e.g. \texttt{MPI\_Recv}). The second one performs computation bursts,
combined  in  between   with  group  communications,  blocking  sends,
immediate  receives and  MPI-wait primitives  are  executed.  Finally,
group  communications interlaced  with computation  and  a significant
synchronization  phase are  performed. From  the performance  point of
view, one can notice a reduction of 2\,\% in the TTS that blocking MPI
mode leads over the polling MPI  policy. On the other hand an averaged
reduction of power dissipation of  2\,\% is attained with the blocking
MPI policy  only in  routines that actually  perform synchronizations,
group  communications  and  MPI-receive/-wait primitives.   Thus,  the
reduction of the power-energy  pair directly depends on the percentage
of time in which blocking MPI  operations are executed. For the 1 hour
simulation traces, an average of 50\,\% of the time the processes were
blocked.  Taking  into account that  the total power  reduction during
blocking phases  is about 5\,\%,  one can compute an  approximation of
the  power savings,  i.e., $50\,\%  \times  5\,\% =  2.5\%$, which  is
consistent  power  reduction  obtained  for  the  full-day  simulation
(around  2\,\%). However,  we  expect higher  power-energy savings  if
future architectures tend proportional computing~\citep{Barroso-2007}.

% \begin{figure*}[htbf]
%   \centering
%   Simulation of 1 hour during midnight (from 23h to 24h) leveraging the polling MPI policy\\
%   \includegraphics[width=0.78\textwidth]{Figs/23_24_blq0_stat.eps}\\
%   Percentage of time for the 1 hour during midnight (from 23h to 24h) leveraging the polling MPI policy\\
%   \includegraphics[width=0.78\textwidth]{Figs/23_24_blq1_stat.eps}\\
%   \caption{blabla}
%   \label{fig:7}
% \end{figure*}