### 2014-05-06 paper_enahpc14/Conclusion.tex
updates latest version

We  have   presented  a  methodology  for   comparing  performance  of
\cosmoart,  a  regional  weather  forecast  model  augmented  for  the
interactions of  reactive gases and aerosol  particles.  The resulting
benchmarks illustrate  that the  best time-to-solution does  not imply
the best energy-to-solution: an Intel Sandybridge (2.6 GHz) system has
lower  time-to-solution but  higher energy-to-solution  than  an Intel
Ivybridge  (2.2  GHz) system,  although  the  two  metrics are  indeed
strongly correlated.  On the  other hand, the usage of energy-friendly
MPI waiting  techniques analysed with  our power-performance framework
over  a  series  of runs  of  \cosmoart  on  \tinto and  power  traces
demonstrate that it  is possible to reduce both  power dissipation and
energy-to-solution   while  maintaining   (or  even   decreasing)  the
time-to-solution. The resulting profiles indicate that simple changes,
such as making use of the  blocking MPI policy rather than the polling
is  possible to  slightly  reduce both  power  dissipation and  energy
consumption (by 2\,\%).

This  reproducible  benchmark provides  a  baseline  for ongoing  work
package within  the EU-funded Exa2Green to  minimise power consumption
of \cosmoart. Profiling  has given us insight into  the most expensive
code  components,  which are  now  being  altered  to utilise  revised
algorithms.  The  results of these  optimisations will be  reported in
future publications.