### 2014-06-16 paper_enahpc14/Sec4:Experimental-results.tex
updates paper according to page proofs

In this section, we present our measurement methodology along with the
metrics  considered  and  discuss about  the  performance-power-energy
trade-off of the model system on the different platforms stated.

\subsection{Measurement methodology}
\label{subsec:4.1}

We  approach  the  assessment  of  the energy  footprint  and  overall
performance    of    \cosmoart    with    two    important    metrics:
\textit{time-to-solution} (TTS) and \textit{energy-to-solution} (ETS).
TTS refers to  the total wall clock time  of the application execution
time. ETS  is the amount of  energy spent to  achieve results.  Energy
consumption is  assessed by sampling the power  during execution which
is then averaged and multiplied  by the TTS to determine ETS. Whenever
possible, multiple  production runs  were performed to  illustrate the
reproducibility  of   the  baseline,  and   quantify  the  significant
uncertainties in  the power measurement, as dictated  by the available
technology.

For  the   experiments  on  \tinto  platform,  we   also  analyze  the
contribution of the MPI library  to the energy consumption and how the
use  of   different  message-passing  progression   engine  modes  can
potentially  render energy  savings. Specifically,  the  OpenMPI 1.6.5
library,  features  two   operation  policies,  \emph{aggressive}  and
\emph{degraded},   which   can  be   selected   before  \cosmoart   is
launched. In the \emph{aggressive}  (by default) policy, normally used
in exactly-  and under-subscribed modes  (i.e., the number  of running
processes is equal to or less than the number of available cores), MPI
processes  will  never voluntarily  give  up  the  processor to  other
processes. With  some network transports, OpenMPI  processes will spin
in  tight   loops  attempting   to  make  message   passing  progress,
effectively causing  other processes  to not get  any CPU  cycles (and
therefore  never make  any progress).   This usually  yields  the best
performance but  a higher CPU utilization,  possibly at the  cost of a
higher energy  usage.  On the other hand,  the \emph{degraded} policy,
suitable for the oversubscribed mode (i.e., more processes are running
than total  cores available), MPI processes will  frequently yield the
processor  to  its  peers,  thereby  allowing all  processes  to  make
progress.

In our experiments with  \cosmoart only the exactly-subscribed mode is
used.  Therefore, \cosmoart processes  running in  the \emph{degraded}
policy are  constantly checking for  the completion of an  event (e.g.
send or receive) and  repeatedly calling to \texttt{sched\_yield()} to
be picked  again by  the Linux scheduler.  Although one can  expect no
benefits  from  using the  \emph{degraded}  policy  compared with  the
\emph{aggressive} one, notable differences  can be observed in the CPU
utilization.  In  a separated  test with the  \emph{aggressive} policy
enabled, we noticed that a blocked MPI process waiting for an incoming
message utilized 100\,\% of CPU, however when the \emph{degraded} mode
was  set,  the  CPU  utilization  was  only  $50\,\%$,  approximately.
Although neither of both policies avoid busy-waiting loops in blocking
routines, the use  of the \emph{degraded} policy is as  much as we can
do in OpenMPI to reduce the CPU utilization, and thus save energy.

% polling configuration
% (the default mode) ``MPI''  continually polls the network interface to
% check for  the completion  of an event  (e.g. send or  receive).  This
% usually yields  low latency  but high CPU  utilization. Thus,  one can
% expect that  this mode attains  the best performance, possibly  at the
% cost a higher energy usage. In  the blocking mode, the CPU is yield to
% other     processes/threads    if     there     are    no     incoming
% messages~\cite{Castillo-2012}.

\subsection{Time-power-energy analysis}
\label{subsec:4.2}

Figures~\ref{fig:1} and \ref{fig:2}  account respectively for \monch's
Isola  E1 Rack 2  and \pilat'  Isola HD  total power  measurements for
1-day or 2-day simulations. On \monch, the 1-day simulation was issued
only twice due  to usage restrictions.  As time  resolution was set to
one averaged update  every 5 minutes for power  sampling, the averaged
power consumption was computed by considering 6 values for each single
run.   On \pilat, the  1-day simulation  was issued  four times  and a
2-day run  only once.  Similarly,  the averaged power  consumption was
computed  by considering  4 values  for each  single 1-day  run  and 9
values  for the  2-days  run. Corresponding  results  are gathered  in
Table~\ref{tab:1}.

\begin{figure}[ht]
  \centering
  \includegraphics[width=1\columnwidth]{Figs/NRJ_benchmark_Monch.eps}
  \caption{Isola E1 Rack 2 total power consumption of \monch.}
  \label{fig:1}
\end{figure}

\begin{figure}[ht]
  \centering
  \includegraphics[width=1\columnwidth]{Figs/NRJ_benchmark_Pilatus.eps}
  \caption{Isola HD total power consumption of \pilat.}
  \label{fig:2}
\end{figure}

\begin{table}[ht]
  \centering
  \caption{Averaged power consumption (W) of the platforms.}
  \label{tab:1}
  \begin{tabular}{cccc}
    \hline\noalign{\smallskip}     \textbf{\scriptsize{PILATUS}}     &
    \textbf{\scriptsize{MONCH}}  &  \textbf{\scriptsize{TINTORRUM}}  &
    \textbf{\scriptsize{TINTORRUM}}           \\          &          &
    \textbf{\scriptsize{(\emph{Aggressive})}} & \textbf{\scriptsize{(\emph{Degraded})}}
    \\    \noalign{\smallskip}\hline\noalign{\smallskip}   18035.0   &
    12622.5 & 3713.6 & 3651.8 \\ \noalign{\smallskip}\hline
  \end{tabular}
\end{table}

In  Figure~\ref{fig:3}, we  compare both  TTS (right  Y-axis)  and ETS
(left Y-axis)  metrics on both  platforms.  As can be  observed, Intel
Sandy Bridge EP  processors of \pilat outperform the  Intel Ivy Bridge
EP processors of \monch, being roughly 130\,\% faster.
% The reason
%for that is two-fold: \emph{i}) they have higher clock frequency than Ivy
%Bridge  (2.6\,GHz  against  2.2\,GHz),   and  \emph{ii}) % they  aim  at
% computing   speed  regardless   to  be   energy  efficient.
In  our  experiments,   \monch  showed  the  best  energy-to-solution,
reducing the energy consumption of \pilat by approximately 7\,\%.

Averaged TTS and ETS results over 10 repetitions of a 1-day simulation
on \tinto  are shown in Figure~\ref{fig:4}.  The corresponding results
for the averaged total power  are also shown in Table~\ref{tab:1}.  As
expected, the average total power consumption by \cosmoart while using
the \emph{degraded} MPI policy is  about 2\,\% less than that consumed
by the \emph{aggressive} one. This is  due to the reduction of the CPU
utilization  (of about  50\,\%)  while using  the \emph{degraded}  MPI
policy, in contrast with the \emph{aggressive} one, that always yields
to 100\,\% of  CPU load while waiting for  incoming messages. From the
point of  view of the execution  time there is a  slight reduction for
the \emph{degraded} MPI  policy. One can expect that  the costs of the
context switches  produced in the \emph{degraded} MPI  policy can lead
to  an increase  of  the execution  time,  however, applications  with
extensive computing,  as e.g.  \cosmoart, may  show better performance
with the \emph{degraded}  policy. This effect works in  favor with our
results, since  the energy  consumed is also  reduced (2\,\%)  just by
changing the  behavior of the  OpenMPI library to  the \emph{degraded}
policy,   i.e,  with   the  \texttt{--mca   mpi\_yield\_when\_idle  1}
parameter set in the \texttt{mpirun} command.

\begin{figure}[ht]
  \centering
  \includegraphics[width=1\columnwidth]{Figs/Time_E2S_COSMO-ART-0.eps}
  \caption{Time-to-solution and  energy-to-solution comparison between
    \pilat and \monch for a 1-day simulation.}
  \label{fig:3}
\end{figure}

\begin{figure}[ht]
  \centering
  \includegraphics[width=1\columnwidth]{Figs/Time_E2S_COSMO-ART-1.eps}
  \vspace{-1cm}
  \caption{Mean time-to-solution and  energy-to-solution on \tinto for
    a 1-day simulation using the \emph{aggressive} and \emph{degraded}
    MPI  policies. The  95\,\% confidence  intervals of  Student's $t$
    distribution (10 samples) clearly indicates an improvement for the
    \emph{degraded} policy in both TTS and ETS.}
  \label{fig:4}
\end{figure}

\begin{figure*}[ht]
  \centering
  \hspace{0.8cm} \scalebox{0.62}{\input{Figs/24hour.pstex_t}}
  \caption{1-day  simulation power  trace using  the \emph{aggressive}
    and \emph{degraded} MPI policies on \tinto.}
  \label{fig:5}
\end{figure*}

\begin{figure*}[ht]
  \centering
  \scriptsize
  1 hour simulation during midday (11h--12h) using the \emph{aggressive} MPI policy \\
  \scalebox{0.5}{\input{Figs/11_12_blq0_figure.pstex_t}} \\
  1 hour simulation during midday (11h--12h) using the \emph{degraded} MPI policy \\
  \scalebox{0.5}{\input{Figs/11_12_blq1_figure.pstex_t}} \\
  1 hour simulation during midnight (23h--24h) using the \emph{aggressive} MPI policy \\
  \scalebox{0.5}{\input{Figs/23_24_blq0_figure.pstex_t}} \\
  1 hour simulation during midnight (23h--24h) using the \emph{degraded} MPI policy \\
  \scalebox{0.5}{\input{Figs/23_24_blq1_figure.pstex_t}} \\
  \begin{tabular}{rlp{-0.2cm}rlp{-0.2cm}rlp{-0.2cm}rlp{-0.2cm}rl}
    & &  & &  & &  \\[-0.15cm]
    Computation     & \multicolumn{1}{>{\columncolor[RGB]{  0,  , 255}}p{0.4cm}}{} & & 
    Block. send     & \multicolumn{1}{>{\columncolor[RGB]{255,  0,174}}p{0.4cm}}{} & & 
    Synchronization & \multicolumn{1}{>{\columncolor[RGB]{179,  0,  0}}p{0.4cm}}{} & & 
    Wait/Wait all   & \multicolumn{1}{>{\columncolor[RGB]{235,255,  0}}p{0.4cm}}{} \\
    & &  & &  & &  \\[-0.15cm]
    Immediate recv. & \multicolumn{1}{>{\columncolor[RGB]{100,100,177}}p{0.4cm}}{} & &
    Waiting a msg.  & \multicolumn{1}{>{\columncolor[RGB]{255,  0  ,0}}p{0.4cm}}{} & & 
    Group comm.     & \multicolumn{1}{>{\columncolor[RGB]{255,144, 26}}p{0.4cm}}{} & &
    Others          & \multicolumn{1}{>{\columncolor[RGB]{192,224,  0}}p{0.4cm}}{} \\
    %I/O             & \multicolumn{1}{>{\columncolor[RGB]{172,174, 41}}p{0.4cm}}{} & &
    %
    %Idle            & \multicolumn{1}{>{\columncolor[RGB]{117,195,255}}p{0.4cm}}{} \\
  \end{tabular}
  \caption{Power-performance   traces  during   midday   and  midnight
    leveraging the \emph{aggressive} and \emph{degraded} MPI policies on \tinto.}
  \label{fig:6}
\end{figure*}

\subsection{Power-performance profiling and tracing}
\label{subsec:4.3}

A full tracing  experiment is conducted on \tinto  in order to capture
an overall power profile at a much finer resolution using the attached
PDUs in each  node. First we run \cosmoart for  a 1-day simulation and
analyze the  power consumption.  Second we  correlate performance with
power  traces  of  two  1-hour  time frames  during  the  day:  midday
(11h--12h) and midnight (23h--24h). All the experiments on \tinto were
performed using 192 cores with  12 MPI processes per node, i.e., using
an exactly-subscribed MPI mode.

In order to obtain an initial overview of the power consumption of the
system  model,  we  obtained  power  profiles  leveraging  our  \pmlib
power-measurement    framework   over    a    1-day   simulation    of
\cosmoart.  These profiles, produced  with both  \emph{aggressive} and
\emph{degraded} MPI policies,  are depicted in Figure~\ref{fig:5}. The
first  observation for the  \emph{aggressive} MPI  policy is  that the
first and last  hours of the day consume slightly  less power than the
midday  hours.   We  relate  these  variations due  to  the  different
chemical   reactions,  aerosols,   radiations  and   clouds  formation
processes, computed by ART, taking place during daylight hours but not
during  night.  In this  case  the  peak  power consumption  is  about
3821\,W. Looking  at the power  profile using the  \emph{degraded} MPI
policy, we  observe that the 1-hour  pattern is repeated  along the 24
hours, favoring  the reduction of power  during night hours  in a more
noticeable way than that  with the \emph{aggressive} policy.  The peak
power  consumption  during day  hours  while  the \emph{degraded}  MPI
policy  is leveraged  is about  3768\,W,  which means  a reduction  of
around  50\,W with  respect to  the \emph{aggressive}  mode.  Also, we
notice systematic  power drops for each simulated  hour.  We associate
them  to periods  in which  the  processes were  waiting for  incoming
messages  or  synchronizations,  thus  yielding  the CPU  to  a  lower
usage. Our final observation is the reduction of 2\,\% in the TTS when
the \emph{degraded} policy is used.

To  gain  more  insights   about  the  power-performance  behavior  of
\cosmoart, we  obtained power-per\-for\-man\-ce traces to  allow us to
correlate  the type  of  tasks  being executed  with  the total  power
consumption  along  the   time  (see  Figure~\ref{fig:6}).   For  this
experiment we limit  the simulation time to 1  hour because: \emph{i})
the  pattern of  the computation  performed by  \cosmoart  is repeated
hourly and, \emph{ii}) the weight  of a 1-day simulation trace can not
be easily handled using our  machines and memory available (each trace
file was  about 45\,GB).   Thus, we shrink  the simulations to  1 hour
during midday (from 11h to 12h) and during midnight (from 23h to 24h),
in which  \cosmoart perform  different internal computations.   At the
same  time   we  leverage  the   two  already  stated   MPI  policies:
\emph{aggressive} and \emph{degraded}.

All  the  traces  in   Figure~\ref{fig:6}  are  divided  in  different
computation  phases. In first  phase all  the processes  perform group
communications      and       execute      MPI-receive      primitives
(e.g. \texttt{MPI\_Recv}). The second one performs computation bursts,
combined  in  between   with  group  communications,  blocking  sends,
immediate   receives   and   MPI-wait  primitives.    Finally,   group
communications   interlaced  with   computation   and  a   significant
synchronization  phase are  performed. From  the performance  point of
view,  one  can notice  a  reduction  of 2\,\%  in  the  TTS that  the
\emph{degraded} MPI  policy leads  over the \emph{aggressive}  one. On
the other hand,  a reduction of power consumption  of 2\,\% in average
is attained  with the \emph{degraded}  policy, only for  routines that
actually  perform  synchronizations,   group  communications  and  MPI
receive and wait primitives.   Thus, the reduction of the power-energy
pair directly depends on the  percentage of time in which blocking MPI
operations are executed. For  the 1-hour simulation traces, an average
of  50\,\%  of  the  time  the processes  were  waiting  for  incoming
messages.  Taking  into account that the total  power reduction during
blocking phases  is about 5\,\%,  one can compute an  approximation of
the  power savings,  i.e., $50\,\%  \times  5\,\% =  2.5\%$, which  is
consistent power  reduction obtained for the  1-day simulation (around
2\,\%).  However, we  expect higher  power/energy savings  if upcoming
versions  of OpenMPI  support full  blocking policies,  as well  as if
future architectures tend proportional computing~\citep{Barroso-2007}.

% \begin{figure*}[htbf]
%   \centering
%   Simulation of 1 hour during midnight (from 23h to 24h) leveraging the polling MPI policy\\
%   \includegraphics[width=0.78\textwidth]{Figs/23_24_blq0_stat.eps}\\
%   Percentage of time for the 1 hour during midnight (from 23h to 24h) leveraging the polling MPI policy\\
%   \includegraphics[width=0.78\textwidth]{Figs/23_24_blq1_stat.eps}\\
%   \caption{blabla}
%   \label{fig:7}
% \end{figure*}